---
title: "Project_4"
author: "EricaWan"
date: "2022-11-30"
output: html_document
---
 

# load libraries / install packages

```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(tidymodels)
library(janitor)
library(skimr)
library(kableExtra)
library(GGally)
library(kableExtra) 
library(vip)        
library(fastshap)   
library(MASS)
library(ISLR)
library(tree)
library(dplyr)
library(ggplot2)
library(imputeMissings)
library(readxl)
library(writexl)
#install.packages("factoextra")
library(factoextra)
library(rpart.plot)
```


# Read data 
```{r, warning=FALSE, message=FALSE}
donationn <- read_csv("C:/Users/16179/Downloads/my_R_Project/Donations.csv")%>% clean_names()
head(donationn) 
skim(donationn)

donor <- read_csv("C:/Users/16179/Downloads/my_R_Project/DonorMerge_Final.csv")%>% clean_names()
head(donor)
skim(donor)

```


#Inner join two data sets
#Remove unique identifers
#skim 
#Remove variables with more than 10% missing values
#Remove categorical variables with too many levels of data
```{r, warning=FALSE, message=FALSE}
donation <- donationn%>%
  inner_join(donor, by=c("projectid" = "projectid")) %>%
  dplyr::select(-projectid, -teacher_acctid, -schoolid, -school_ncesid, -donationid, -donor_acctid, -donation_timestamp,-donor_zip, 
                -donor_city, -donor_state, -donation_message,  -school_metro,-school_city, -school_state, -school_district, -school_county, -date_posted, -primary_focus_subject, 
                -great_messages_proportion, -teacher_referred_count, -non_teacher_referred_count, -one_non_teacher_referred_donor_g, -secondary_focus_subject, -secondary_focus_area, -school_zip )


skim(donation)
head(donation)



```

#Exploratory analysis:
#1. pct of is_excitings 
```{r,warning=FALSE, message=FALSE}
donation %>%
  ggplot(aes(x=is_exciting)) +
  geom_histogram(stat="count") +
  labs(title = "How many project is exciting")

donation %>%
  group_by(is_exciting) %>%
  summarize(n=n()) %>%
  ungroup() %>%
  mutate(pct = n/sum(n))

```


#scale data for further clustering
```{r, warning=FALSE, message=FALSE}

dontation_s <- donation

#fulfillment_labor_materials
dontation_s$fulfillment_labor_materials[is.na(dontation_s$fulfillment_labor_materials)]<- median(dontation_s$fulfillment_labor_materials, na.rm =TRUE)
#students_reached
dontation_s$students_reached[is.na(dontation_s$students_reached)]<- median(dontation_s$students_reached, na.rm =TRUE)

#primary_focus_area
dontation_s$primary_focus_area[is.na(dontation_s$primary_focus_area)]<- mode(dontation_s$primary_focus_area)
#resource_type
dontation_s$resource_type[is.na(dontation_s$resource_type)]<- mode(dontation_s$resource_type)
#grade_level
dontation_s$grade_level[is.na(dontation_s$grade_level)]<- mode(dontation_s$grade_level)


#logical variables
#is_teacher_acct
dontation_s$is_teacher_acct_T <- ifelse(dontation_s$is_teacher_acct == 'TRUE', 1, 0)
dontation_s$is_teacher_acct_F <- ifelse(dontation_s$is_teacher_acct == 'FALSE', 1, 0)

#donation_included_optional_support
dontation_s$donation_included_optional_support_T <- ifelse(dontation_s$donation_included_optional_support == 'TRUE', 1, 0)
dontation_s$donation_included_optional_support_F <- ifelse(dontation_s$donation_included_optional_support == 'FALSE', 1, 0)

#payment_included_acct_credit
dontation_s$payment_included_acct_credit_T <- ifelse(dontation_s$payment_included_acct_credit == 'TRUE', 1, 0)
dontation_s$payment_included_acct_creditt_F <- ifelse(dontation_s$payment_included_acct_credit == 'FALSE', 1, 0)


#payment_included_campaign_gift_card
dontation_s$payment_included_campaign_gift_card_T <- ifelse(dontation_s$payment_included_campaign_gift_card == 'TRUE', 1, 0)
dontation_s$payment_included_campaign_gift_card_F <- ifelse(dontation_s$payment_included_campaign_gift_card == 'FALSE', 1, 0)

#payment_included_web_purchased_gift_card
dontation_s$payment_included_web_purchased_gift_card_T <- ifelse(dontation_s$payment_included_web_purchased_gift_card == 'TRUE', 1, 0)
dontation_s$payment_included_web_purchased_gift_card_F <- ifelse(dontation_s$payment_included_web_purchased_gift_card == 'FALSE', 1, 0)

#payment_was_promo_matched
dontation_s$payment_was_promo_matched_T <- ifelse(dontation_s$payment_was_promo_matched == 'TRUE', 1, 0)
dontation_s$payment_was_promo_matched_F <- ifelse(dontation_s$payment_was_promo_matched == 'FALSE', 1, 0)

#via_giving_page
dontation_s$via_giving_page_T <- ifelse(dontation_s$via_giving_page == 'TRUE', 1, 0)
dontation_s$via_giving_page_F <- ifelse(dontation_s$via_giving_page == 'FALSE', 1, 0)


#for_honoree
dontation_s$for_honoree_T <- ifelse(dontation_s$for_honoree == 'TRUE', 1, 0)
dontation_s$for_honoree_F <- ifelse(dontation_s$for_honoree == 'FALSE', 1, 0)



#school_charter
dontation_s$school_charter_T <- ifelse(dontation_s$school_charter == 'TRUE', 1, 0)
dontation_s$school_charter_F <- ifelse(dontation_s$school_charter == 'FALSE', 1, 0)

#school_magnet
dontation_s$school_magnet_T <- ifelse(dontation_s$school_magnet == 'TRUE', 1, 0)
dontation_s$school_magnet_F <- ifelse(dontation_s$school_magnet == 'FALSE', 1, 0)

#school_year_round
dontation_s$school_year_round_T <- ifelse(dontation_s$school_year_round == 'TRUE', 1, 0)
dontation_s$school_year_round_F <- ifelse(dontation_s$school_year_round == 'FALSE', 1, 0)

#school_nlns
dontation_s$school_nlns_T <- ifelse(dontation_s$school_nlns == 'TRUE', 1, 0)
dontation_s$school_nlns_F <- ifelse(dontation_s$school_nlns == 'FALSE', 1, 0)

#school_kipp
dontation_s$school_kipp_T <- ifelse(dontation_s$school_kipp == 'TRUE', 1, 0)
dontation_s$school_kipp_F <- ifelse(dontation_s$school_kipp == 'FALSE', 1, 0)

#school_charter_ready_promise
dontation_s$school_charter_ready_promise_T <- ifelse(dontation_s$school_charter_ready_promise == 'TRUE', 1, 0)
dontation_s$school_charter_ready_promise_F <- ifelse(dontation_s$school_charter_ready_promise == 'FALSE', 1, 0)

#teacher_teach_for_america
dontation_s$teacher_teach_for_america_T <- ifelse(dontation_s$teacher_teach_for_america == 'TRUE', 1, 0)
dontation_s$teacher_teach_for_america_F <- ifelse(dontation_s$teacher_teach_for_america == 'FALSE', 1, 0)

#teacher_ny_teaching_fellow
dontation_s$teacher_ny_teaching_fellow_T <- ifelse(dontation_s$teacher_ny_teaching_fellow == 'TRUE', 1, 0)
dontation_s$teacher_ny_teaching_fellow_F <- ifelse(dontation_s$teacher_ny_teaching_fellow == 'FALSE', 1, 0)

#eligible_double_your_impact_matc
dontation_s$eligible_double_your_impact_matc_T <- ifelse(dontation_s$eligible_double_your_impact_matc == 'TRUE', 1, 0)
dontation_s$eligible_double_your_impact_matc_F <- ifelse(dontation_s$eligible_double_your_impact_matc == 'FALSE', 1, 0)

#eligible_almost_home_match
dontation_s$eligible_almost_home_match_T <- ifelse(dontation_s$eligible_almost_home_match == 'TRUE', 1, 0)
dontation_s$eligible_almost_home_match_F <- ifelse(dontation_s$eligible_almost_home_match == 'FALSE', 1, 0)


# create dummy variables for categorical variable class
#dollar_amount

dontation_s$dollar_amount_10_to_100 <- ifelse(dontation_s$dollar_amount == '10_to_100', 1, 0)
dontation_s$dollar_amount_under_10 <- ifelse(dontation_s$dollar_amount == 'under_10', 1, 0)
dontation_s$dollar_amount_100_and_up <- ifelse(dontation_s$dollar_amount == '100_and_up', 1, 0)


#payment method
dontation_s$payment_method_no_cash_received <- ifelse(dontation_s$payment_method  == 'no_cash_received', 1, 0)
dontation_s$payment_method_paypal <- ifelse(dontation_s$payment_method == 'paypal', 1, 0)
dontation_s$payment_method_creditcard <- ifelse(dontation_s$payment_method == 'creditcard', 1, 0)
dontation_s$payment_method_amazon <- ifelse(dontation_s$payment_method == 'amazon', 1, 0)
dontation_s$payment_method_promo_code_match <- ifelse(dontation_s$payment_method  == 'promo_code_match', 1, 0)
dontation_s$payment_method_check <- ifelse(dontation_s$payment_method == 'check', 1, 0)
dontation_s$payment_method_almost_home_match <- ifelse(dontation_s$payment_method == 'almost_home_match', 1, 0)
dontation_s$payment_method_double_your_impact_match <- ifelse(dontation_s$payment_method == 'double_your_impact_match', 1, 0)


#teacher prefix
dontation_s$teacher_prefix_Mrs. <- ifelse(dontation_s$teacher_prefix == 'Mrs.', 1, 0)
dontation_s$teacher_prefix_Ms. <- ifelse(dontation_s$teacher_prefix == 'Ms.', 1, 0)
dontation_s$teacher_prefix_Mr. <- ifelse(dontation_s$teacher_prefix == 'Mr.', 1, 0)


#primary_focus_area
dontation_s$primary_focus_area <- ifelse(dontation_s$primary_focus_area  == 'Literacy & Language', 1, 0)
dontation_s$primary_focus_area <- ifelse(dontation_s$primary_focus_area == 'History & Civics', 1, 0)
dontation_s$primary_focus_area <- ifelse(dontation_s$primary_focus_area == 'Math & Science', 1, 0)
dontation_s$primary_focus_area <- ifelse(dontation_s$primary_focus_area == 'Music & The Arts', 1, 0)
dontation_s$primary_focus_area <- ifelse(dontation_s$primary_focus_area  == 'Health & Sports', 1, 0)
dontation_s$primary_focus_area <- ifelse(dontation_s$primary_focus_area == 'Special Needs', 1, 0)
dontation_s$primary_focus_area <- ifelse(dontation_s$primary_focus_area == 'Applied Learning', 1, 0)

#resource_type
dontation_s$resource_type_Technology <- ifelse(dontation_s$resource_type  == 'Technology', 1, 0)
dontation_s$resource_type_Supplies <- ifelse(dontation_s$resource_type == 'Supplies', 1, 0)
dontation_s$resource_type_Other <- ifelse(dontation_s$resource_type == 'Other', 1, 0)
dontation_s$resource_type_Books <- ifelse(dontation_s$resource_type == 'Books', 1, 0)
dontation_s$resource_type_Trips <- ifelse(dontation_s$resource_type  == 'Trips', 1, 0)
dontation_s$resource_type_Visitors <- ifelse(dontation_s$resource_type == 'Visitors', 1, 0)

#poverty_level
dontation_s$poverty_level <- ifelse(dontation_s$poverty_level  == 'highest poverty', 1, 0)
dontation_s$poverty_level <- ifelse(dontation_s$poverty_level == '	
moderate poverty', 1, 0)
dontation_s$poverty_level <- ifelse(dontation_s$poverty_level == '	
low poverty
', 1, 0)
dontation_s$poverty_level <- ifelse(dontation_s$poverty_level == '	
low poverty
', 1, 0)


#grade_level
dontation_s$grade_level <- ifelse(dontation_s$grade_level  == '	
Grades 3-5', 1, 0)
dontation_s$grade_level <- ifelse(dontation_s$grade_level == '	
Grades 9-12', 1, 0)
dontation_s$grade_level <- ifelse(dontation_s$grade_level == '	
Grades 6-8', 1, 0)
dontation_s$grade_level <- ifelse(dontation_s$grade_level == '	
Grades PreK-2', 1, 0)



#standardize numeric variables

dontation_s$donation_to_project_s <- scale(dontation_s$donation_to_project)
dontation_s$donation_optional_support_s <- scale(dontation_s$donation_optional_support)

dontation_s$donation_total_s <- scale(dontation_s$donation_total)
dontation_s$school_latitude_s <- scale(dontation_s$	school_latitude)
dontation_s$school_longitude_s <- scale(dontation_s$ school_longitude)
dontation_s$fulfillment_labor_materials_s <- scale(dontation_s$	fulfillment_labor_materials)
dontation_s$total_price_excluding_optional_s_s <- scale(dontation_s$total_price_excluding_optional_s)
dontation_s$total_price_including_optional_s_s <- scale(dontation_s$total_price_including_optional_s)
dontation_s$students_reached_s <- scale(dontation_s$students_reached)




# remove redundant and rejected variables
donationfin = subset(dontation_s, select= -c(donation_to_project, donation_total, donation_optional_support, school_latitude, school_longitude,  fulfillment_labor_materials, total_price_excluding_optional_s, total_price_including_optional_s, students_reached, is_teacher_acct, donation_included_optional_support, payment_included_acct_credit, payment_included_campaign_gift_card,payment_included_web_purchased_gift_card, payment_was_promo_matched, via_giving_page, for_honoree, school_charter, school_magnet, school_year_round, school_nlns, school_kipp, school_charter_ready_promise, teacher_teach_for_america, teacher_ny_teaching_fellow, eligible_double_your_impact_matc, eligible_almost_home_match, dollar_amount,payment_method,teacher_prefix,primary_focus_area,resource_type,poverty_level, grade_level
)) 

head(donationfin)
skim(donationfin)

```



# build clusters
#I will use 5 as for starting point

```{r, warning=FALSE, message=FALSE}
set.seed(20)

clusters7 <- kmeans(donationfin, 4, iter.max = 200, nstart = 4)
print(clusters7)

# visualize clusters

f<-fviz_cluster(clusters7,donationfin,ellipse.type="norm",geom="point")
dontation_s%>%
  mutate(clusters=clusters7[["cluster"]])

dontation_s%>%
  group_by(clusters7$cluster)%>%
  summarise(mean_dm = mean(donation_to_project_s))
f

```


# explore clusters
```{r, warning=FALSE, message=FALSE}

cluster <- as.factor(clusters7$cluster)

clusters7

#determine which variables are driving the cluster creation

tree.clusters=tree(cluster~.,donationfin)


summary(tree.clusters)
plot(tree.clusters)
text(tree.clusters,pretty=1)
tree.clusters

```



#visual clusters
```{r, warning=FALSE, message=FALSE}

ggplot(dontation_s,aes(x=donation_to_project))+geom_histogram(binwidth=10000)+
  coord_cartesian(xlim=c(0,85000))
ggplot(dontation_s,aes(x=donation_to_project))+geom_histogram(binwidth=10000) + facet_wrap(~clusters7$cluster)+
  coord_cartesian(xlim=c(0,85000))

ggplot(dontation_s,aes(x=donation_total))+geom_histogram(binwidth=10000)+
  coord_cartesian(xlim=c(0,100000))
ggplot(dontation_s,aes(x=donation_total))+geom_histogram(binwidth=10000) + facet_wrap(~clusters7$cluster)+
  coord_cartesian(xlim=c(0,100000))
ggplot(dontation_s,aes(x=donation_optional_support))+geom_histogram(binwidth=1000)+
  coord_cartesian(xlim=c(0,15000))
ggplot(dontation_s,aes(x=donation_optional_support))+geom_histogram(binwidth=10000) + facet_wrap(~clusters7$cluster)+
  coord_cartesian(xlim=c(0,15000))
ggplot(dontation_s,aes(x=school_latitude))+geom_histogram()+
  coord_cartesian(xlim=c(0,68))
ggplot(dontation_s,aes(x=school_latitude))+geom_histogram() + facet_wrap(~clusters7$cluster)+
  coord_cartesian(xlim=c(0,68))



ggplot(dontation_s,aes(x=fulfillment_labor_materials))+geom_histogram()
ggplot(dontation_s,aes(x=fulfillment_labor_materials))+geom_histogram()+facet_wrap(~clusters7$cluster)


ggplot(dontation_s,aes(x=is_teacher_acct))+geom_bar()
ggplot(dontation_s,aes(x=is_teacher_acct))+geom_bar()+facet_wrap(~clusters7$cluster)

ggplot(dontation_s,aes(x=donation_included_optional_support))+geom_bar()
ggplot(dontation_s,aes(x=donation_included_optional_support))+geom_bar()+facet_wrap(~clusters7$cluster)

ggplot(dontation_s,aes(x=payment_included_acct_credit))+geom_bar()
ggplot(dontation_s,aes(x=payment_included_acct_credit))+geom_bar()+facet_wrap(~clusters7$cluster)

ggplot(dontation_s,aes(payment_included_campaign_gift_card))+geom_bar()
ggplot(dontation_s,aes(payment_included_campaign_gift_card))+geom_bar()+facet_wrap(~clusters7$cluster)

ggplot(dontation_s,aes(payment_included_web_purchased_gift_card))+geom_bar()
ggplot(dontation_s,aes(payment_included_web_purchased_gift_card))+geom_bar()+facet_wrap(~clusters7$cluster)

ggplot(dontation_s,aes(payment_was_promo_matched))+geom_bar()
ggplot(dontation_s,aes(payment_was_promo_matched))+geom_bar()+facet_wrap(~clusters7$cluster)


ggplot(dontation_s,aes(via_giving_page))+geom_bar()
ggplot(dontation_s,aes(via_giving_page))+geom_bar()+facet_wrap(~clusters7$cluster)
ggplot(dontation_s,aes(for_honoree))+geom_bar()
ggplot(dontation_s,aes(for_honoree))+geom_bar()+facet_wrap(~clusters7$cluster)
ggplot(dontation_s,aes(school_charter))+geom_bar()
ggplot(dontation_s,aes(school_charter))+geom_bar()+facet_wrap(~clusters7$cluster)
ggplot(dontation_s,aes(school_magnet))+geom_bar()
ggplot(dontation_s,aes(school_magnet))+geom_bar()+facet_wrap(~clusters7$cluster)
ggplot(dontation_s,aes(school_year_round))+geom_bar()
ggplot(dontation_s,aes(school_year_round))+geom_bar()+facet_wrap(~clusters7$cluster)
ggplot(dontation_s,aes(school_nlns))+geom_bar()
ggplot(dontation_s,aes(school_nlns))+geom_bar()+facet_wrap(~clusters7$cluster)
ggplot(dontation_s,aes(school_kipp))+geom_bar()
ggplot(dontation_s,aes(school_kipp))+geom_bar()+facet_wrap(~clusters7$cluster)
ggplot(dontation_s,aes(school_charter_ready_promise))+geom_bar()
ggplot(dontation_s,aes(school_charter_ready_promise))+geom_bar()+facet_wrap(~clusters7$cluster)
ggplot(dontation_s,aes(teacher_teach_for_america))+geom_bar()
ggplot(dontation_s,aes(teacher_teach_for_america))+geom_bar()+facet_wrap(~clusters7$cluster)
ggplot(dontation_s,aes(teacher_ny_teaching_fellow))+geom_bar()
ggplot(dontation_s,aes(teacher_ny_teaching_fellow))+geom_bar()+facet_wrap(~clusters7$cluster)
ggplot(dontation_s,aes(eligible_double_your_impact_matc))+geom_bar()
ggplot(dontation_s,aes(eligible_double_your_impact_matc))+geom_bar()+facet_wrap(~clusters7$cluster)
ggplot(dontation_s,aes(eligible_almost_home_match))+geom_bar()
ggplot(dontation_s,aes(eligible_almost_home_match))+geom_bar()+facet_wrap(~clusters7$cluster)
ggplot(dontation_s,aes(dollar_amount))+geom_bar()
ggplot(dontation_s,aes(dollar_amount))+geom_bar()+facet_wrap(~clusters7$cluster)
ggplot(dontation_s,aes(payment_method))+geom_bar()
ggplot(dontation_s,aes(payment_method))+geom_bar()+facet_wrap(~clusters7$cluster)
ggplot(dontation_s,aes(eligible_almost_home_match))+geom_bar()
ggplot(dontation_s,aes(eligible_almost_home_match))+geom_bar()+facet_wrap(~clusters7$cluster)
ggplot(dontation_s,aes(teacher_prefix))+geom_bar()
ggplot(dontation_s,aes(teacher_prefix))+geom_bar()+facet_wrap(~clusters7$cluster)
ggplot(dontation_s,aes(primary_focus_area))+geom_bar()
ggplot(dontation_s,aes(primary_focus_area))+geom_bar()+facet_wrap(~clusters7$cluster)
ggplot(dontation_s,aes(resource_type))+geom_bar()
ggplot(dontation_s,aes(resource_type))+geom_bar()+facet_wrap(~clusters7$cluster)
ggplot(dontation_s,aes(poverty_level))+geom_bar()
ggplot(dontation_s,aes(poverty_level))+geom_bar()+facet_wrap(~clusters7$cluster)
ggplot(dontation_s,aes(grade_level))+geom_bar()
ggplot(dontation_s,aes(grade_level))+geom_bar()+facet_wrap(~clusters7$cluster)


ggplot(dontation_s,aes(is_exciting))+geom_bar()
ggplot(dontation_s,aes(is_exciting))+geom_bar()+facet_wrap(~clusters7$cluster)
#+ labels(paste0(rounded , "%"))

```

```{r,warning=FALSE, message=FALSE}
res<- dontation_s%>%
  group_by(clusters7$cluster, is_exciting)%>%
   summarize(n=n()) %>%
  ungroup() %>%
  mutate(pct = n/sum(n))
res

```


#convert to factor levels
```{r, warning=FALSE, message=FALSE}
donation %>%
    mutate_if(is.character, factor)%>%
    mutate_if(is.logical, as.factor)%>%
    mutate(is_exciting =  as.factor(is_exciting)) -> donation_prep

head(donation_prep)

```

#first partition: since the data set is too big to run, I select 80% of data set
```{r, warning=FALSE, message=FALSE}
# -- set a random seed for repeatablity 
set.seed(123)

# -- performs our train / donor split 
donation_split_pp <- initial_split(donation_prep, prop = 0.8)

# -- extract the training data 
donation_train_pp <- training(donation_split_pp)

donation_test_pp <- testing(donation_split_pp)

sprintf("Train PCT : %1.2f%%", nrow(donation_train_pp)/ nrow(donation) * 100)
sprintf("Test  PCT : %1.2f%%", nrow(donation_test_pp)/ nrow(donation) * 100)

head(donation_train_pp)

```


## Partition your data 70/30 (train / test split) 
```{r, warning=FALSE, message=FALSE}
# -- set a random seed for repeatablity 
set.seed(123)

# -- performs our train / donor split 
donation_split <- initial_split(donation_train_pp, prop = 0.7)

# -- extract the training data 
donation_train <- training(donation_split)
# -- extract the donor data 
donation_test <- testing(donation_split)

sprintf("Train PCT : %1.2f%%", nrow(donation_train)/ nrow(donation) * 100)
sprintf("Test  PCT : %1.2f%%", nrow(donation_test)/ nrow(donation) * 100)

head(donation_train)
skim(donation_train)
```

#examine relationship between numeric variables 
```{r, warning=FALSE, message=FALSE}
numeric_subset <- na.omit(subset(donation, select = c(donation_to_project, donation_total, donation_optional_support, school_latitude, school_longitude, fulfillment_labor_materials, total_price_excluding_optional_s, total_price_including_optional_s, students_reached, is_teacher_acct, donation_included_optional_support, payment_included_acct_credit)))
skim(numeric_subset)

cormat <- cor(numeric_subset)
round(cormat, 2)

```

#create recipe for model
```{r, warning=FALSE, message=FALSE}
 
donation_recipe <- recipe(is_exciting ~ .
                       , data = donation_train) %>%
  step_impute_mean(all_numeric_predictors()) %>%
  step_impute_mode(all_nominal_predictors())%>%
  step_scale(all_numeric())%>%
  step_novel(all_nominal_predictors())%>%
  step_dummy(all_nominal_predictors(), one_hot = TRUE)%>%
  prep()

donation_recipe

```

## Bake recipe
```{r, warning=FALSE, message=FALSE}
# -- apply the recipe 
bake_train <- bake(donation_recipe, new_data = donation_train)
skim(bake_train)
bake_test  <- bake(donation_recipe, new_data = donation_test)

```



## Interpret with a logistic regression or tree 

#first model
```{r, warning=FALSE, message=FALSE}
logistic_re <-logistic_reg(mode = "classification") %>%
                  set_engine("glm") %>%
                  fit(is_exciting ~ ., data = bake_train)


tidy(logistic_re) %>%
  mutate_at(c("estimate", "std.error", "statistic", "p.value"),round, 4)


```


```{r, warning=FALSE, message=FALSE}
# -- training 
predict(logistic_re, bake_train, type = "prob") %>%
  bind_cols(.,predict(logistic_re, bake_train)) %>%
  bind_cols(.,bake_train) -> scored_train_log

head(scored_train_log)

# -- testing 
predict(logistic_re, bake_test, type = "prob") %>%
  bind_cols(.,predict(logistic_re, bake_test)) %>%
  bind_cols(.,bake_test) -> scored_test_log

head(scored_test_log)
```


```{r, warning=FALSE, message=FALSE}
options(yardstick.event_first = FALSE)

scored_train_log %>% 
  metrics(is_exciting, .pred_TRUE, estimate = .pred_class) %>%
  mutate(part="training") %>%
  bind_rows( scored_train_log %>% 
               metrics(is_exciting, .pred_TRUE, estimate = .pred_class) %>%
               mutate(part="testing") 
  ) 


# -- Variable Importance top 10 features  
logistic_re %>%
  vip(num_features = 10)


# -- ROC Charts 
scored_train_log %>%
  mutate(model = "train") %>%
  bind_rows(scored_test_log %>%
              mutate(model="bake_test")) %>%
  group_by(model) %>%
  roc_curve(is_exciting, .pred_TRUE) %>%
  autoplot()


# -- Confustion Matricies  
scored_train_log %>%
  conf_mat(is_exciting, .pred_class) %>%
  autoplot( type = "heatmap") +
  labs(title="Train Confusion Matrix")

scored_test_log %>%
  conf_mat(is_exciting, .pred_class) %>%
  autoplot( type = "heatmap") +
  labs(title="Test Confusion Matrix")
```



### -  try a diff tree 
```{r, warning=FALSE, message=FALSE}
donation_tree2 <- decision_tree(mode="classification",
                            cost_complexity = 0.0001,
                            tree_depth = 8,
                            min_n = 10) %>%
                  set_engine("rpart") %>%
                  fit(is_exciting ~.
, data=bake_train)



donation_tree2$fit

options(scipen = 0)

rpart.plot(donation_tree2$fit, fallen.leaves = TRUE, roundint=FALSE,cex = 0.4, extra=3)

```

## Prep for Evaluation (donation tree)

```{r, warning=FALSE, message=FALSE}

# -- training 
predict(donation_tree2, bake_train, type = "prob") %>%
  bind_cols(.,predict(donation_tree2, bake_train)) %>%
  bind_cols(.,bake_train) -> scored_train_tree

head(scored_train_tree)

# -- testing 
predict(donation_tree2, bake_test, type = "prob") %>%
  bind_cols(.,predict(donation_tree2, bake_test)) %>%
  bind_cols(.,bake_test) -> scored_test_tree

head(scored_test_tree)
``` 



# Evaluate
```{r, warning=FALSE, message=FALSE}

options(yardstick.event_first = FALSE)
# -- AUC: Train and Test 
scored_train_tree %>% 
  metrics(is_exciting, .pred_TRUE, estimate = .pred_class) %>%
  mutate(part="training") %>%
  bind_rows( scored_test_tree %>% 
               metrics(is_exciting, .pred_TRUE, estimate = .pred_class) %>%
               mutate(part="testing") 
  ) 
  


# -- Variable Importance top 10 features  
donation_tree2 %>%
  vip(num_features = 5)

# -- ROC Charts 
scored_train_tree %>%
  mutate(model = "train") %>%
  bind_rows(scored_test_tree %>%
              mutate(model="test")) %>%
  group_by(model) %>%
  roc_curve(is_exciting, .pred_TRUE) %>%
  autoplot()


# -- Confustion Matricies  
scored_train_tree %>%
  conf_mat(is_exciting, .pred_class) %>%
  autoplot( type = "heatmap") +
  labs(title="Train Confusion Matrix")

scored_test_tree %>%
  conf_mat(is_exciting, .pred_class) %>%
  autoplot( type = "heatmap") +
  labs(title="Test Confusion Matrix")

```






