
#Load libraries
```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(tidymodels)
library(janitor)
library(skimr)
library(kableExtra)
library(GGally)
library(kableExtra) 
library(vip)        # --  tidymodels variable importance
library(fastshap)   # -- shapley values for variable importance 
library(MASS)
```


#Import dataset
```{r, message=FALSE, warning=FALSE}


unidata <-read_csv("C:/Users/16179/Downloads/my_R_Project/churn_training.csv") %>%
  clean_names()

head(unidata) 

```
#Data summary/overview
```{r, message=FALSE, warning=FALSE}

unidata%>%
  skim()

  

```

#Explore enrollment balance
```{r, message=FALSE, warning=FALSE}
unidata%>%
  ggplot(aes(x=churn))+
  geom_histogram(stat="count")+
  labs("Enrollment Count: Yes and No")

pct_exam <- unidata%>%
  group_by(churn)%>%
  summarize(n=n())%>%
  ungroup()%>%
  mutate(pct = n/sum(n))
pct_exam
```

#Explore Boxplots of all numeric and barchart for categorical
```{r, message=FALSE, warning=FALSE}

unidata$churn <- as.factor(unidata$churn)
unidata$phone_model <- as.factor(unidata$phone_model)
unidata$partner <- as.factor(unidata$partner)
unidata$phone_service <- as.factor(unidata$phone_service)
unidata$phone_area_code <- as.factor(unidata$phone_area_code)
unidata$streaming_plan <- as.factor(unidata$streaming_plan)
unidata$multiple_lines <- as.factor(unidata$multiple_lines)
unidata$device_protection <- as.factor(unidata$device_protection)
unidata$billing_state <- as.factor(unidata$billing_state)
unidata$paperless_billing <- as.factor(unidata$paperless_billing)
unidata$payment_method <- as.factor(unidata$payment_method)
unidata$network_speed <- as.factor(unidata$network_speed)
unidata$mobile_hotspot <- as.factor(unidata$mobile_hotspot)
unidata$wifi_calling_text <- as.factor(unidata$wifi_calling_text)
unidata$online_backup <- as.factor(unidata$online_backup)
unidata$gender <- as.factor(unidata$gender)

unidata$contract_code <- as.factor(unidata$contract_code)
unidata$currency_code <- as.factor(unidata$currency_code)
unidata$maling_code <- as.factor(unidata$maling_code)
unidata$email_domain<- as.factor(unidata$email_domain)

```
#commercial household income,campus visit count,self-initiated contact count,total contact count, excurricular interest, mail qualify score, attendence to campus event


#set seed for datasets
```{r, warning=FALSE}
set.seed(999)

uni_split <- initial_split(unidata, prop = 0.7)

uni_train <- training(uni_split)
uni_test <- testing(uni_split)

sprintf("Train PCT : %1.2f%%", nrow(uni_train)/ nrow(unidata) * 100)
sprintf("Test  PCT : %1.2f%%", nrow(uni_test)/ nrow(unidata) * 100)

```

#2. Recipe
#Drop all rejected variables from model and some variables with more than 20% missing values 
```{r, warning=FALSE}
uni_recipe <- recipe(churn~ email_domain+ monthly_minutes + total_billed+ customer_service_calls+ mobile_hotspot+ late_payments+
multiple_lines+ currency_code+ phone_model+ streaming_minutes + prev_balance+ payment_method+  phone_area_code + paperless_billing+ number_phones+ streaming_plan, 
data= uni_train)%>%
  
  step_impute_median(all_numeric_predictors()) %>%
  step_unknown(all_nominal_predictors()) %>%
  step_scale(all_numeric_predictors()) %>%
  step_dummy(all_nominal_predictors())%>%

  prep()

uni_recipe

   
 
```


#apply recipe to two data sets
```{r, warning=FALSE}
bake_train <- bake(uni_recipe, new_data = uni_train)
  skim(bake_train)
bake_test  <- bake(uni_recipe, new_data = uni_test)
  skim(bake_test)

```


```{r, warning=FALSE}
logistic_uni <-logistic_reg(mode = "classification") %>%
                  set_engine("glm") %>%
                  fit(churn ~ ., data = bake_train)


tidy(logistic_uni) %>%
  mutate_at(c("estimate", "std.error", "statistic", "p.value"),round, 4)


```


```{r, warning=FALSE}
# -- training 
predict(logistic_uni, bake_train, type = "prob") %>%
  bind_cols(.,predict(logistic_uni, bake_train)) %>%
  bind_cols(.,bake_train) -> scored_train_uni

head(scored_train_uni)

# -- testing 
predict(logistic_uni, bake_test, type = "prob") %>%
  bind_cols(.,predict(logistic_uni, bake_test)) %>%
  bind_cols(.,bake_test) -> scored_test_uni

head(scored_test_uni)
```


```{r, warning=FALSE}
options(yardstick.event_first=FALSE)

scored_train_uni %>% 
  metrics(churn, .pred_1, estimate = .pred_class) %>%
  mutate(part="training") %>%
  bind_rows(scored_test_uni %>% 
               metrics(churn, .pred_1, estimate = .pred_class) %>%
               mutate(part="testing") 
  ) 


# -- Variable Importance top 10 features  
logistic_uni %>%
  vip(num_features = 50)


# -- ROC Charts 
scored_train_uni %>%
  mutate(model = "train") %>%
  bind_rows(scored_test_uni %>%
              mutate(model="test")) %>%
  group_by(model) %>%
  roc_curve(churn, .pred_1) %>%
  autoplot()


# -- Confustion Matricies  
scored_train_uni %>%
  conf_mat(churn, .pred_class) %>%
  autoplot( type = "heatmap") +
  labs(title="Train Confusion Matrix")

scored_test_uni %>%
  conf_mat(churn, .pred_class) %>%
  autoplot( type = "heatmap") +
  labs(title="Test Confusion Matrix")
```

```{r, warning=FALSE}
steplog <- glm(churn ~ ., data = bake_train, family=binomial(link="logit"))
step <- stepAIC(steplog, direction="both")
summary(step)


```

```{r, warning=FALSE}
## -- Use tidymodel framework to fit and evaulate reduced model



uni_steprecipe <- recipe(enroll ~ campus_visit+premiere+stuemail+instate+interest+init_span+int1rat+int2rat+hscrat, data = uni_train) %>%
  prep()

uni_steprecipe


# -- apply new recipe 
bake_steptrain <- bake(uni_steprecipe, new_data = uni_train)
bake_steptest  <- bake(uni_steprecipe, new_data = uni_test)

logistic_step1 <-logistic_reg(mode = "classification") %>%
                  set_engine("glm") %>%
                  fit(enroll ~ ., data = bake_steptrain)


## -- check out your parameter estimates ... 
tidy(logistic_step1) %>%
  mutate_at(c("estimate", "std.error", "statistic", "p.value"),round, 4)


```

```{r, warning=FALSE}

# -- training predictions from stepwise model
predict(logistic_step1, bake_steptrain, type = "prob") %>%
  bind_cols(.,predict(logistic_step1, bake_steptrain)) %>%
  bind_cols(.,bake_steptrain) -> scored_train_step1

head(scored_train_step1)

# -- testing predictions from stepwise model
predict(logistic_step1, bake_steptest , type = "prob") %>%
  bind_cols(.,predict(logistic_step1, bake_steptest )) %>%
  bind_cols(.,bake_steptest ) -> scored_test_step1

head(scored_test_step1)

```
```{r, warning=FALSE}
# -- Evaluate Stepwise Model
# -- AUC: Train and Test 
options(yardstick.event_first=FALSE)
scored_train_step1 %>% 
  metrics(enroll, .pred_1, estimate = .pred_class) %>%
  mutate(part="training") %>%
  bind_rows( scored_test_step1 %>% 
               metrics(enroll, .pred_1, estimate = .pred_class) %>%
               mutate(part="testing") 
  )



# -- ROC Charts 
scored_train_step1 %>%
  mutate(model = "train") %>%
  bind_rows(scored_test_step1 %>%
              mutate(model="test")) %>%
  group_by(model) %>%
  roc_curve(enroll, .pred_1) %>%
  autoplot()

```
```{r}
# -- Confustion Matricies  
scored_train_step1 %>%
  conf_mat(enroll, .pred_class) %>%
  autoplot( type = "heatmap") +
  labs(title="Train Confusion Matrix")

scored_test_step1 %>%
  conf_mat(enroll, .pred_class) %>%
  autoplot( type = "heatmap") +
  labs(title="Test Confusion Matrix")






```
```{r, warning=FALSE}




uni_step_2_recipe <- recipe(enroll ~ campus_visit+premiere+interest+instate+init_span+int1rat+int2rat+hscrat, data = uni_train) %>%
  prep()

uni_step_2_recipe


# -- apply new recipe 
bake_steptrain_2 <- bake(uni_step_2_recipe, new_data = uni_train)
bake_steptest_2  <- bake(uni_step_2_recipe, new_data = uni_test)

logistic_step2 <-logistic_reg(mode = "classification") %>%
                  set_engine("glm") %>%
                  fit(enroll ~ ., data = bake_steptrain_2)


## -- check out your parameter estimates ... 
tidy(logistic_step2) %>%
  mutate_at(c("estimate", "std.error", "statistic", "p.value"),round, 4)


```

```{r, warning=FALSE}

predict(logistic_step2, bake_steptrain_2, type = "prob") %>%
  bind_cols(.,predict(logistic_step2, bake_steptrain_2)) %>%
  bind_cols(.,bake_steptrain_2) -> scored_train_step2

head(scored_train_step2)

predict(logistic_step2, bake_steptest_2 , type = "prob") %>%
  bind_cols(.,predict(logistic_step2, bake_steptest_2 )) %>%
  bind_cols(.,bake_steptest_2 ) -> scored_test_step2

head(scored_test_step2)

```


```{r, warning=FALSE}
# -- Evaluate Model

options(yardstick.event_first=FALSE)

scored_train_step2 %>% 
  metrics(enroll, .pred_1, estimate = .pred_class) %>%
  mutate(part="training") %>%
  bind_rows( scored_test_step2 %>% 
               metrics(enroll, .pred_1, estimate = .pred_class) %>%
               mutate(part="testing") 
  )



# -- ROC Charts 
scored_train_step2 %>%
  mutate(model = "train") %>%
  bind_rows(scored_test_step2 %>%
              mutate(model="test")) %>%
  group_by(model) %>%
  roc_curve(enroll, .pred_1) %>%
  autoplot()

```


```{r}
# -- Confusion Matrices  
scored_train_step2 %>%
  conf_mat(enroll, .pred_class) %>%
  autoplot( type = "heatmap") +
  labs(title="Train Confusion Matrix")

scored_test_step2 %>%
  conf_mat(enroll, .pred_class) %>%
  autoplot( type = "heatmap") +
  labs(title="Test Confusion Matrix")


```
